---
title: "Getting started with cgvR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with cgvR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

cgvR renders 3D graphs with a Vulkan backend. This vignette walks through the
main APIs on small built-in examples: opening a viewer, uploading a graph,
laying it out, controlling the camera, highlighting a path, and recording a
short video.

Code that opens a real window (`cgv_run`) is shown with `eval = FALSE` so the
vignette builds in a headless environment. Run those chunks interactively to
see the result.

```{r}
library(cgvR)
```

---

## 1. A first graph

Start with a tiny graph: 8 nodes on the corners of a cube, edges along the
cube's edges. Positions are explicit, no layout needed.

```{r}
nodes <- 1:8
edges <- cbind(
  c(1,2,3,4, 5,6,7,8, 1,2,3,4),
  c(2,3,4,1, 6,7,8,5, 5,6,7,8)
)

# corner coordinates of a unit cube, scaled
pos <- matrix(c(
  -1,-1,-1,   1,-1,-1,   1, 1,-1,  -1, 1,-1,
  -1,-1, 1,   1,-1, 1,   1, 1, 1,  -1, 1, 1
), ncol = 3, byrow = TRUE) * 5
```

```{r, eval = FALSE}
v <- cgv_viewer(800, 600, "cube")
cgv_background(v, "black")
cgv_set_graph(v, nodes, edges,
              positions   = pos,
              node_values = as.double(seq_len(8)),
              node_sizes  = rep(20, 8))
cgv_camera(v, position = c(15, 12, 18), target = c(0, 0, 0))
cgv_run(v)
```

`node_values` is mapped through a colormap (viridis by default — see `cmap` in
`?cgv_set_graph` for plasma / inferno / magma).

---

## 2. Force-directed layout

For graphs without natural coordinates use `cgv_layout_fr` (pure-R
Fruchterman-Reingold) or `cgv_layout_fr_bh` (Barnes-Hut, O(n log n) in C —
prefer it for thousands of nodes).

```{r}
set.seed(1)
n <- 60L
# random tree + a few extra edges
ef <- 1L; et <- integer(0)
for (i in 2:n) { ef <- c(ef, sample.int(i - 1, 1)); et <- c(et, i) }
ef <- c(ef, sample.int(n, 20)); et <- c(et, sample.int(n, 20))
edges <- cbind(ef[seq_len(min(length(ef), length(et)))],
               et[seq_len(min(length(ef), length(et)))])

pos <- cgv_layout_fr(n, edges, n_iter = 200L, seed = 42L)
str(pos)
```

```{r, eval = FALSE}
v <- cgv_viewer(1000, 700, "FR layout")
cgv_set_graph(v, seq_len(n), edges,
              positions   = pos,
              node_values = as.double(seq_len(n)),
              node_sizes  = rep(10, n))
cgv_camera(v, position = c(20, 16, 24), target = c(0, 0, 0))
cgv_run(v)
```

Same call, larger graph, faster algorithm:

```{r, eval = FALSE}
pos <- cgv_layout_fr_bh(5000L, edges_big, n_iter = 200L, seed = 1L)
```

---

## 3. Highlighting a path

`cgv_highlight_path` overlays a path on top of the existing graph. Pass any
sequence of node ids; consecutive ids get a thick coloured edge between them.

```{r, eval = FALSE}
v <- cgv_viewer(1000, 700, "path demo")
cgv_set_graph(v, seq_len(n), edges, positions = pos,
              node_values = as.double(seq_len(n)),
              node_sizes  = rep(8, n))

cgv_highlight_path(v, c(1, 5, 17, 42), color = "#FF2200",
                   node_scale = 2.0, edge_width = 4.0)

# remove the highlight again:
# cgv_clear_path(v)

cgv_run(v)
```

---

## 4. Camera control

Three modes are available — `fly` (WASD + mouse), `orbit`, `arcball`. Switch at
runtime with `cgv_camera_mode`. Two helpers animate the camera:

* `cgv_fly_to(v, node_id, duration)` — fly to a single node.
* `cgv_fly_path(v, node_ids, duration)` — animate along a Catmull-Rom spline
  through several waypoints.

```{r, eval = FALSE}
v <- cgv_viewer(1000, 700, "camera demo")
cgv_set_graph(v, seq_len(n), edges, positions = pos,
              node_sizes = rep(8, n))

cgv_camera_mode(v, "orbit")
cgv_fly_path(v, c(1, 17, 42, 5, 1), duration = 6.0)
cgv_run(v)
```

---

## 5. Recording a video

`cgv_record_start` / `cgv_record_stop` pipe frames to `ffmpeg` (must be on
`PATH`). In offscreen mode you control the number of frames explicitly via
`cgv_run(v, n_frames = ...)`, which is convenient for scripted rendering.

```{r, eval = FALSE}
v <- cgv_viewer(1280, 720, "record demo", offscreen = TRUE)
cgv_set_graph(v, seq_len(n), edges, positions = pos,
              node_sizes = rep(8, n))
cgv_camera(v, position = c(20, 16, 24), target = c(0, 0, 0))

cgv_record_start(v, "demo.mp4", fps = 30)
cgv_run(v, n_frames = 90L)        # 3 seconds at 30 fps
cgv_record_stop(v)
```

---

## 6. Headless mode

`offscreen = TRUE` skips the GLFW window. Useful for tests, CI, and anywhere a
display is unavailable.

```{r, eval = FALSE}
v <- cgv_viewer(640, 480, offscreen = TRUE)
cgv_set_graph(v, nodes, edges, positions = pos)
cgv_run(v, n_frames = 1L)
cgv_close(v)
```

---

## Where to next

The `inst/examples/` directory ships ready-to-run scripts:

* `demo_small_graph.R` — random 100-node graph with FR layout.
* `demo_cycles_bh.R` — TopSpin cycles with Barnes-Hut FR.
* `demo_tictactoe.R` — Tic-Tac-Toe state graph; node colour = move number.
* `demo_record.R`, `demo_cycles_bh_record.R` — video recording end-to-end.

Run them with `Rscript inst/examples/<file>.R` from the package root.
